Water, High-Altitude Condensates, and Methane Depletion in the Atmosphere of the Warm Neptune WASP-107\lowercase{b}

### 2017-08-17 ms.tex
v3

\author{Michael Line}
\affiliation{Arizona State University}
\nocollaboration

@@

\author{Caroline Morley}
\altaffiliation{Sagan Fellow}
\affiliation{Harvard-Smithsonian Center for Astrophysics}
\nocollaboration

\author{Kevin Stevenson}
\affiliation{Space Telescope Science Institute}
\nocollaboration

\begin{abstract}
The Neptune-mass exoplanet WASP-107b is an exciting target for atmosphere characterization. It has an unusually large atmospheric scale height and a small, bright host star, raising the possibility of precise constraints on its current nature and formation history.  We report the first atmospheric study of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.  We determined the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius. We detect water at $6.5\,\sigma$ confidence and set an upper limit on the atmospheric metallicity of $30\times$ solar. We find that the methane abundance is depleted relative to expectations for a solar abundance pattern (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.  The water features are smaller than expected for a cloudless atmosphere, crossing less than one scale height, and requires a thick condensate layer at high altitudes (0.1 - 3 mbar). We find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at these pressures.  Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres. The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities. 
\end{abstract}

@@

WASP-107b also has a relatively low equilibrium temperature (780 K) compared to most other exoplanets that are amenable to atmosphere characterization.  This results in a distinct atmospheric chemistry from other well-studied systems: at low temperatures, the dominant molecular reservoir for carbon transitions from carbon monoxide to methane \citep{moses13}.  Spectral features from both water and methane are accessible with current observing facilities, enabling a spectroscopic estimate of the carbon-to-oxygen ratio (C/O). Previous measurements of C/O have been challenging because they rely on broadband photometry or narrow wavelength coverage \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.

In addition, WASP-107b is one of the best targets discovered to date for atmosphere characterization. Thanks to its large atmospheric scale height and small, bright host star, the expected signal-to-noise for the transmission spectrum is comparable to the best studied benchmarks in the field (e.g. HD\,209458b, HD\,189733b). The planet is also a feasible target for thermal emission measurements, with an expected secondary eclipse depth of $\sim0.1$\% in \emph{Spitzer}'s $4.5\,\mu$m channel. In this paper we report the first atmosphere characterization of WASP-107b: a near-infrared transmission spectrum measured with the \emph{Hubble Space Telescope} (\emph{HST}; Program GO 14915, PI L. Kreidberg).

\section{Observations}
We observed a single transit of WASP-107b with \emph{HST}'s Wide Field Camera 3 (WFC3) instrument on UT 5-6 June 2017.  The transit observation consisted of five \emph{HST} orbits. At the beginning of each 96-minute orbit, we took an image of the target with the F130N filter (exposure time = 4.2 s). This direct image is used for wavelength calibration. For the remainder of the target visibility period (about 45 minutes), we obtained time series spectra with the G141 grism, which provides low-resolution spectroscopy over the wavelength range $1.1 - 1.7\,\mu$m.  We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool\footnote{\url{https://github.com/spacetelescope/PandExo\_HST}}.  As is standard for WFC3 observations of bright targets, we used the spatial scanning observing mode, which slews the telescope in the spatial direction over the course of an exposure. The scan rate was 0.12 arcseconds/sec.

\section{Data Reduction}
We reduced the data with the custom pipeline described in \cite{kreidberg14a}, which we summarize briefly here. For each exposure, we extracted the spectrum from each up-the-ramp sample (or ``stripe") separately using the optimal extraction algorithm of \cite{horne86}. The stripe spectra were then summed to create the final spectrum. For each stripe, the extraction box was 80 pixels high and centered on the stripe's midpoint in the spatial direction. We corrected the spectra for the changing dispersion solution over the length of the spatial scan and small drifts over time ($<0.1$ pixel).  To subtract the background, we took the median of sky pixels that were uncontaminated by the target spectrum (rows $5-250$, columns $5-15$). 
%dispersion solution: 46 Angstroms/pixel. max drift: 3 angstroms (from diagnostics.txt)

@@

\subsection{White Light Curve}
To create the raw white light curve, we summed each spectrum over the 181 pixels in the spectral trace.  The white light curve has systematic trends that are typical for WFC3 observations \citep{zhou17}: the flux increases increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend. The largest ramp occurs in the initial orbit (orbit zero), so we only fit data from orbits one through four in our analysis, following common practice.  We fit the light curve with the analytic model of the form $F_\mathrm{white}(t) = S_\mathrm{white}(t)\times T_\mathrm{white}(t)$, where $S_\mathrm{white}$ is a systematics model and $T_\mathrm{white}$ is a transit model. We used the same systematics model as \cite{kreidberg15b}, Equation 1.  To model the transit, we used the \texttt{batman} package \citep{kreidberg15a}.  The transit model parameters are the orbital period $p$, time of inferior conjunction $t_0$, transit depth $r_p/r_s$, ratio of semi-major axis to stellar radius $a/r_s$, orbital inclination $i$, and the quadratic stellar limb darkening parameters $u_1$ and $u_2$.

\begin{figure*}
\includegraphics[width = \textwidth]{Figures/lc.pdf}
\caption{Broadband and spectrophotometric transit light curves for WASP-107b measured with \textit{HST}/WFC3 compared to best fit models (left) and residuals from the best fit (right). Annotations indicate the central wavelength and root-mean-square (rms) residuals for each light curve. A possible star-spot crossing feature is shaded in gray in the upper right; our systematic error correction removes this feature from the spectroscopic light curves.} 
\label{fig:lc}
\end{figure*}

@@

\subsubsection{Final Fit}
In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the precise estimates from the Kepler light curve \citep{dai17}.  These values are consistent with our best fit when we allowed them to vary freely. We also fixed the limb darkening parameters on predictions from a PHOENIX model for a star with effective temperature 4300 K, calculated with the \texttt{limb-darkening} package from \cite{espinoza15}.  The remaining free parameters were $t_0$, $r_p/r_s$, and the systematics parameters for the visit-long and orbit-long trends.
%The model values are consistent with what we obtain from fitting for the limb darkening coefficients as free parameters.

@@

The best fit light curves have a median $\chi^2_\nu$ value of 1.16.  To ensure that we did not underestimate the uncertainty on the transit depths, we rescaled the photometric uncertainties for all spectroscopic channels such that the $\chi^2_\nu$ values are unity. We performed an MCMC fit to the light curves with \texttt{emcee}.  For each light curve we ran a fit with 50 walkers and 1000 steps per walker, and tested for convergence as we did for the white light curve. The median transit depths and $1\,\sigma$ uncertainties are given in Table\,\ref{tab:transit_depths}.

We explored several alternative choices for the spectroscopic light curve fits, but found that none of them made a significant difference in the transmission spectrum. For example, we fit the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths (differing by just $0.3\sigma$ on average), except with a small constant offset due to the uncorrected star-spot crossing feature. This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval (see \S\,\ref{sec:retrieval}). We also fit for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model values. We found that the fitted limb darkening coefficients are consistent with the model predictions, so we opted to fix the coefficients in our final analysis because it improves the precision on the transit depths by about 10\%. We also checked that the uncertainty in the stellar parameters does not significantly affect the PHOENIX model predictions at the level of precision of our data.  We also performed an independent data reduction and fit using K. Stevenson's pipeline and found consistent results (well within $1\,\sigma$).

\subsubsection{Transit Depths Redder than 1.62 $\mu$m}
The red edge of the transmission spectrum is of interest for our analysis because methane is expected to be the dominant absorber over water at wavelengths greater than $1.6\,\mu$m. Unfortunately, we find that the our reddest spectroscopic light curves ($>1.62\mu$m) are too poor in quality to robustly measure the transit depth. The residuals exhibit correlated noise and the fits have $\chi^2_\nu>2$. The transit depths are also sensitive to which method we use to fit for instrument systematics.  We therefore opt not to report transit depths redder than the 1.62 $\mu$m channel.

\begin{figure*}
\includegraphics[width = \textwidth]{Figures/spectrum.pdf}
\caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals). The features at $1.15$ and $1.4$ $\mu$m are due to water absorption. The right-hand axis indicates the normalized atmospheric scale height assuming a $10\times$ solar composition.}
\label{fig:spectrum}
\end{figure*}

@@

\subsection{Atmospheric Metallicity from Interior Structure Modeling}
\label{sec:interior}
Given WASP-107b's unusually low density, we quantitatively explored the range of atmospheric metallicities that are consistent with the observed mass and radius using the structure evolution modeling of \cite{thorngren16}.  These models assume a thermally inert heavy-element core with a convective envelope of additively mixed H/He \citep{saumon95} and heavy-element impurities.  The heavy elements were a 50-50 rock-ice mix from ANEOS \citep{thompson90}.  We evolved the planets in time using the atmospheric models of \cite{fortney07}.  The results are sensitive to assumptions about the stellar age, which is uncertain \citep[either $0.6\pm0.2$ to $8.3\pm4.3$ Gyr depending on model assumptions;][]{mocnik17}. We therefore ran two models, with uniform age priors of either $0.2-1.0$ or $1.0-13.8$ Gyr.  We used the published mass and radius estimates \citep[$0.12\pm0.01\,M_\mathrm{J}$, $0.94\pm0.02$;][]{anderson17}.  %A more precise mass estimate is in preparation (B. Benneke, priv.\,comm.); however, we find that our results are not significantly affected when we vary the assumed mass by $\pm3\,\sigma$ from the published value because the dominant uncertainties are the stellar age and core mass.

Based on these assumptions, we fit for envelope metallicity and core mass using an MCMC with uniform priors on both parameters.  The MCMC burned in for $10^3$ steps and then collected $4\times10^6$ samples, which we thinned to $10^5$.  The resulting envelope metal mass fractions were converted to metallicities by assuming the mean molecular weight of the metals was 18 (the value for water), using the approach of \cite{fortney13}. Figure\,\ref{fig:metal_prior} shows the results.  We find an upper limit on the metallicity of $30\times$ solar for the young stellar age range.  Higher metallicity envelopes are not allowed because they decrease the planet's radius below the observed value.  For the older age, the upper limit is even lower (FIXME), because planets cool and contract as they age \citep{fortney08}.  The inferred metallicity and core mass are inversely proportional, as expected (the more massive the core, the lighter the envelope must be to maintain the observed radius). The upper limit on the core mass is FIXME.

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/metalByAge.pdf}
\caption{Estimated envelope metallicity for a given core mass based on interior structure modeling. The red shading corresponds to the $3\sigma$ confidence interval for a young planet ($0.2-1$ Gyr); the blue is for an older age ($1-14$ Gyr).}
\label{fig:metal_prior}
\end{figure}

\subsection{Retrieval}
\label{sec:retrieval}
We also inferred the composition of the atmosphere directly from the transmission spectrum using the CHIMERA chemically-consistent retrieval \citep{line13a, kreidberg15b}.   Briefly, CHIMERA solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.  We parameterized atmospheric composition with metallicity and carbon-to-oxygen ratio under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon94} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.    We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} from the pre-tabulated line-by-line cross section database described in \cite{freedman14}. The transmission forward model is coupled with the PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.

@@

%There is one $3.4\sigma$ outlier data point (at 1.34 $\mu$m), so we explored a few retrieval scenarios with additional complexity.  We tested the addition of cloud patchiness (as described by \citealt{line16}) and a quench pressure for nitrogen and carbon species \citep[to account for transport-induced disequilibrium chemistry, e.g.][]{morley17}.  These models did not significantly improve the fit quality, and the main retrieval results were unchanged. We also tested the effect of varying the planet's surface gravity within the published FIXME $\sigma$ range, and found that our results were again minimally changed.

The nominal retrieval results are shown in Figure\,\ref{fig:retrieval}.  The best fit model provides a good fit to the data ($\chi^2_\nu = 1.2$).  The metallicity distribution spans the full range allowed by our priors, with some preference for larger values. The cloud top pressure is estimated to be $0.01 - 3$ mbar at $1\,\sigma$ confidence. The retrieved irradiation temperature is $525 - 820$ K (1 $\sigma$ confidence).  We find that the C/O value is less than solar (0.54) at $2.7\,\sigma$ confidence.  We tested the detection significance for water by removing water opacity from the nominal model. The Bayesian evidence favors the inclusion of water at $6.5\,\sigma$ confidence.

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/retrieval.pdf}
\caption{Retrieved distributions for (a) metallicity, (b) carbon-to-oxygen ratio, (c) irradiation temperature, and (d) cloud-top pressure in bars. Dashed vertical lines show the median and $\pm1\,\sigma$ confidence interval. Solar metallicity and C/O values are indicated with solid vertical lines.}  \label{fig:retrieval}
\end{figure}

\subsubsection{Methane Abundance}
We ran two additional retrievals to explore whether methane depletion is the underlying cause of the inferred low C/O.  Even though water is the dominant absorber in the WFC3 bandpass, there is enough methane opacity to change the shape of the spectral features at a detectable level for high precision data.  In one test, we assumed chemical equilibrium but excluded methane opacity. This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$), resolving the tension with the solar value.

We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium. The $3\,\sigma$ upper limit on the methane abundance is $1.2\times10^{-3}$. We compared this to the expected methane abundance for the temperature and pressure sensed at the terminator.  Assuming  solar C/O and $3\,\times$ solar metallicity (the median value for the chemical equilbrium retrieval), the methane abundance is predicted to be $1.3\times10^{-3}$, inconsistent with our estimated value.  Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern.

\subsection{Condensate Properties}
In addition to the atmospheric retrieval, we also performed forward modeling of physically motivated, self-consistent clouds and hazes using the methods described in \cite{fortney08, morley15}.  We model clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS; see \citealt{morley12}), for a range of metallicites from $1-50\times$ solar. We vary the cloud sedimentation efficiency from 0.1 to 3 (highly lofted, small particulate clouds). None of the models produce sufficiently low amplitude spectral features to match the observed spectrum. The results are sensitive to surface gravity, but even when we increase the gravity by FIXME $\sigma$ the clouds still form too low.

We also model an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere. We predict the abundance of hydrocarbon haze precursors from previous photochemical models for GJ 436b \citep{line11, morley17}. With the most efficient haze production (100\% of precursors form haze) and particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations. However, these photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis, and we found in the previous section that methane is depleted in the atmosphere of WASP-107b.  Even if the precursors were present, it is unclear whether the 100\% haze production efficiency is realistic. Further modeling of the condensates is needed for a satisfactory explanation of the data.

%We do not consider ``sloped hazes" at this time as they are unlikely to influence the results.
%We also switch off the quench pressure parameters as they are unlikely to strongly influence the equilibrium abundance profiles of the spectrally prominent, nearly constant-with-altitude abundance, species like CH$_4$, H$_2$O, and NH$_3$ over this temperature range \citep[e.g.][]{line11, moses13}.  
\section{Discussion and Conclusions} \label{sec:discuss}
We analyzed the atmospheric composition of WASP-107b based on retrieval of its near-infrared transmission spectrum and interior structure models of the planet's mass and radius.  Key results from this analysis include:

>>>>

\author{Michael R. Line}
\affiliation{Arizona State University}
\nocollaboration

@@

\author{Caroline V. Morley}
\altaffiliation{Sagan Fellow}
%\affiliation{Harvard-Smithsonian Center for Astrophysics}
\affiliation{Harvard-Smithsonian Center for Astrophysics\
60 Garden St.\\
Cambridge, MA 02138}
\nocollaboration

\author{Kevin B. Stevenson}
\affiliation{Space Telescope Science Institute}
\nocollaboration

\begin{abstract}
The Neptune-mass exoplanet WASP-107b is an exciting target for atmosphere characterization. It has an unusually large atmospheric scale height and a small, bright host star, raising the possibility of precise constraints on its current nature and formation history.  We report the first atmospheric study of WASP-107b, a Hubble Space Telescope measurement of its near-infrared transmission spectrum.  We determined the planet's composition with two techniques: atmospheric retrieval based on the transmission spectrum and interior structure modeling based on the observed mass and radius. We detect water at $6.5\,\sigma$ confidence and set an upper limit on the atmospheric metallicity of $30\times$ solar. We find that the methane abundance is depleted relative to expectations for a solar abundance pattern (at $3\,\sigma$ confidence), suggesting a low carbon-to-oxygen ratio or high internal heat flux.  The water features are smaller than expected for a cloudless atmosphere, crossing less than one scale height, and require a thick condensate layer at high altitudes (0.1 - 3 mbar). We find that it is challenging for physically motivated cloud and haze models to produce opaque condensates at these pressures.  Taken together, these findings serve as an illustration of the diversity and complexity of exoplanet atmospheres. The community can look forward to more such results with the high precision and wide spectral coverage afforded by future observing facilities. 
\end{abstract}

@@

WASP-107b also has a relatively low equilibrium temperature compared to most other exoplanets that are amenable to atmosphere characterization (780 K, assuming zero albedo).  This results in a distinct atmospheric chemistry from other well-studied systems: at low temperatures, the dominant molecular reservoir for carbon transitions from carbon monoxide to methane \citep{moses13}.  Spectral features from both water and methane are accessible with current observing facilities, enabling a spectroscopic estimate of the carbon-to-oxygen ratio (C/O). Previous measurements of C/O have been challenging because they rely on broadband photometry or narrow wavelength coverage \citep[e.g.][]{madhusudhan11, line14, benneke15, kreidberg15b}.

In addition, WASP-107b is one of the best targets discovered to date for atmosphere characterization. Thanks to its large atmospheric scale height and small, bright host star, the expected signal-to-noise for the transmission spectrum is comparable to the best studied benchmarks in the field (e.g. HD\,209458b).  In this paper we report the first atmosphere characterization of WASP-107b: a near-infrared transmission spectrum measured with the \emph{Hubble Space Telescope} (\emph{HST}; Program GO 14915, PI L. Kreidberg).

\section{Observations and Data Reduction}
We observed a single transit of WASP-107b with \emph{HST}'s Wide Field Camera 3 (WFC3) instrument on UT 5-6 June 2017.  The transit observation consisted of five \emph{HST} orbits. At the beginning of each 96-minute orbit, we took an image of the target with the F130N filter (exposure time = 4.2 s). This direct image is used for wavelength calibration. For the remainder of the target visibility period (about 45 minutes), we obtained time series spectra with the G141 grism, which provides low-resolution spectroscopy over the wavelength range $1.1 - 1.7\,\mu$m.  We used the NSAMP=6, SPARS\_25 readout mode (exposure time = 112 s) to optimize the efficiency of the observations, as determined by the \texttt{PandExo\_HST} planning tool\footnote{\url{https://github.com/spacetelescope/PandExo\_HST}}.  As is standard for WFC3 observations of bright targets, we used the spatial scanning observing mode, which slews the telescope in the spatial direction over the course of an exposure. The scan rate was 0.12 arcseconds/sec.

We reduced the data with the custom pipeline described in \cite{kreidberg14a}, which we summarize briefly here. For each exposure, we extracted the spectrum from each up-the-ramp sample (or ``stripe") separately using the optimal extraction algorithm of \cite{horne86}. The stripe spectra were then summed to create the final spectrum. For each stripe, the extraction box was 80 pixels high and centered on the stripe's midpoint in the spatial direction. We corrected the spectra for the changing dispersion solution over the length of the spatial scan and small drifts over time ($<0.1$ pixel).  To subtract the background, we took the median of sky pixels that were uncontaminated by the target spectrum (rows $5-250$, columns $5-15$). 
%dispersion solution: 46 Angstroms/pixel. max drift: 3 angstroms (from diagnostics.txt)

@@

\subsection{White Light Curve}
To create the raw white light curve, we summed each spectrum over the 181 pixels in the spectral trace.  The white light curve has systematic trends that are typical for WFC3 observations \citep{zhou17}: the flux increases asymptotically over each orbit (the ``ramp" effect) and there is a visit-long linear trend. The largest ramp occurs in the initial orbit (orbit zero), so we only fit data from orbits one through four in our analysis, following common practice.  We fit the light curve with the analytic model of the form $F_\mathrm{white}(t) = S_\mathrm{white}(t)\times T_\mathrm{white}(t)$, where $S_\mathrm{white}$ is a systematics model and $T_\mathrm{white}$ is a transit model. We used the same systematics model as \cite{kreidberg15b}.  We modelled the transit with the \texttt{batman} package \citep{kreidberg15a}.  The model parameters are the orbital period $p$, time of inferior conjunction $t_0$, transit depth $r_p/r_s$, ratio of semi-major axis to stellar radius $a/r_s$, orbital inclination $i$, and the quadratic stellar limb darkening parameters $u_1$ and $u_2$.

\begin{figure*}
\includegraphics[width = \textwidth]{Figures/lc.pdf}
\caption{Left: Broadband and spectrophotometric transit light curves for WASP-107b compared to best fit models. Right: residuals from the best fit. Annotations indicate the central wavelength and root-mean-square (rms) residuals. A star-spot crossing feature is shaded in gray in the upper right; our systematic error correction removes this feature from the spectroscopic light curves.} 
\label{fig:lc}
\end{figure*}

@@

\subsubsection{Final Fit}
In our final fit, we fixed the transit parameters $a/r_s$, $i$, $p$ on the precise estimates from the Kepler light curve \citep{dai17}.  We also fixed the limb darkening parameters on predictions from a PHOENIX model for a star with effective temperature 4300 K, calculated with the \texttt{limb-darkening} package from \cite{espinoza15}.  We checked that the values we fixed were consistent with our estimates when we allowed them to vary freely.  We also checked that the uncertainty in the stellar parameters does not significantly affect the PHOENIX model predictions at the level of precision of our data.  The remaining free parameters in the fit were $t_0$, $r_p/r_s$, and the systematics parameters for the visit-long and orbit-long trends.

@@

The best fit light curves have a median $\chi^2_\nu$ value of 1.16.  We conservatively rescaled the photometric uncertainties for all spectroscopic channels such that the $\chi^2_\nu$ values are unity. We performed an MCMC fit to the light curves with \texttt{emcee}.  For each light curve we ran a fit with 50 walkers and 1000 steps per walker, and tested for convergence as we did for the white light curve. The median transit depths and $1\,\sigma$ uncertainties are given in Table\,\ref{tab:transit_depths}.

We explored several alternative choices for the spectroscopic light curve fits, but found that none of them made a significant difference in the transmission spectrum. For example, we fit the spectroscopic light curves with the same analytic model we used for the white light curve and obtained nearly identical relative transit depths. 
%(differing by just $0.3\sigma$ on average), except with a small constant offset due to the uncorrected star-spot crossing feature. This offset does not affect our final analysis because it is marginalized in the atmospheric retrieval (see \S\,\ref{sec:retrieval}). 
We also fit for a linear limb darkening parameter rather than fixing the limb darkening on the PHOENIX model values. We found that the fitted limb darkening coefficients are consistent with the model predictions, so we opted to fix the coefficients in our final analysis because it improves the precision on the transit depths by about 10\%.  We also performed an independent data reduction and fit using K. Stevenson's pipeline and found consistent results (well within $1\,\sigma$).

%\subsubsection{Transit Depths Redder than 1.62 $\mu$m}
%The red edge of the transmission spectrum is of interest for our analysis because methane is expected to be the dominant absorber over water at wavelengths greater than $1.6\,\mu$m. Unfortunately, we find that the our reddest spectroscopic light curves ($>1.62\mu$m) are too poor in quality to robustly measure the transit depth. The residuals exhibit correlated noise and the fits have $\chi^2_\nu>2$. The transit depths are also sensitive to which method we use to fit for instrument systematics.  We therefore opt not to report transit depths redder than the 1.62 $\mu$m channel.

\begin{figure*}
\includegraphics[width = \textwidth]{Figures/spectrum.pdf}
\caption{The transmission spectrum of WASP-107b (points with 1\,$\sigma$ error bars) compared to retrieved models (blue line with shaded $1$ and $2\sigma$ confidence intervals). The features at $1.15$ and $1.4$ $\mu$m are due to water absorption. The right-hand axis indicates the normalized atmospheric scale height assuming a $10\times$ solar composition. The water feature crosses less than one scale height, indicating that condensates are present at high altitude.} 
\label{fig:spectrum}
\end{figure*}

@@

\subsection{Atmospheric Metallicity from Interior Structure Modeling}
\label{sec:interior}
Given WASP-107b's unusually low density, we quantitatively explored the range of atmospheric metallicities that are consistent with the observed mass and radius using the structure evolution modeling of \cite{thorngren16}.  These models assume a thermally inert heavy-element core with a convective envelope of additively mixed H/He \citep{saumon95} and heavy-element impurities.  The heavy elements were a 50-50 rock-ice mix. We evolved the planets in time using the atmospheric models of \cite{fortney07}.  The results are sensitive to assumptions about the stellar age, which is uncertain \citep[either $0.6\pm0.2$ to $8.3\pm4.3$ Gyr depending on model assumptions;][]{mocnik17}. We therefore ran two models, with uniform age priors of either $0.2-1.0$ or $1.0-13.8$ Gyr.  We used the published mass and radius estimates \citep[$0.12\pm0.01\,M_\mathrm{J}$, $0.94\pm0.02$;][]{anderson17}.  %A more precise mass estimate is in preparation (B. Benneke, priv.\,comm.); however, we find that our results are not significantly affected when we vary the assumed mass by $\pm3\,\sigma$ from the published value because the dominant uncertainties are the stellar age and core mass.

Based on these assumptions, we fit for envelope metallicity and core mass using an MCMC with uniform priors on both parameters.  The MCMC burned in for $10^3$ steps and then collected $4\times10^6$ samples.  The envelope metal mass fractions were converted to metallicities assuming the mean molecular weight of the metals was 18 (the value for water), using the approach of \cite{fortney13}. Figure\,\ref{fig:metal_prior} shows the results.  We find an upper limit on the metallicity of $30\times$ solar for the young stellar age range.  Higher metallicity envelopes are not allowed because they decrease the planet's radius below the observed value.  For the older age, the upper limit is even lower ($20\times$ solar), because planets cool and contract as they age \citep{fortney08}.  These are conservative upper limits, but realistically, the planet probably formed with a core. Assuming a $5\,M_\oplus$ core, the upper limit on metallicity is 20 (10) $M_\oplus$ for the young (old) age.

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/metalByAge.pdf}
\caption{Estimated envelope metallicity for a given core mass based on interior structure modeling. The red shading corresponds to the $1\sigma$ confidence interval for a young planet ($0.2-1$ Gyr); the blue is for an older age ($1-14$ Gyr).}
\label{fig:metal_prior}
\end{figure}

\subsection{Retrieval}
\label{sec:retrieval}
We also inferred the composition of the atmosphere directly from the transmission spectrum using the CHIMERA chemically-consistent retrieval \citep{line13a}.   Briefly, CHIMERA solves the transmission geometry problem using the equations in \cite{brown01, tinetti12}.  We parameterized atmospheric composition with metallicity and carbon-to-oxygen ratio under the assumption of thermochemical equilibrium using the NASA CEA routine \citep{gordon94} to compute the molecular abundances for H$_2$, He, H$_2$O, CH$_4$, CO, CO$_2$, NH$_3$, H$_2$S, Na, K, HCN, C$_2$H$_2$, TiO, VO, and FeH.    We updated the transmission model to use correlated-K opacities \citep{lacis91, molliere15, amundsen16} from the pre-tabulated line-by-line cross section database described in \cite{freedman14}. The transmission forward model is coupled with the PyMultiNest tool \citep{buchner16} to solve the parameter estimation and model selection problems.

@@

%There is one $3.4\sigma$ outlier data point (at 1.34 $\mu$m), so we explored a few retrieval scenarios with additional complexity.

The nominal retrieval results are shown in Figure\,\ref{fig:retrieval}.  The best fit model provides a good fit to the data ($\chi^2_\nu = 1.2$).  The metallicity distribution spans the full range allowed by our priors, with preference for larger values. The cloud top pressure is estimated to be $0.01 - 3$ mbar at $1\,\sigma$ confidence. The retrieved irradiation temperature is $525 - 820$ K (1 $\sigma$ confidence).  We find that the C/O value is less than solar (0.54) at $2.7\,\sigma$ confidence.  We tested the detection significance for water by removing water opacity from the nominal model. The Bayesian evidence favors the inclusion of water at $6.5\,\sigma$ confidence.

We explored a few retrieval scenarios with additional complexity, including the addition of cloud patchiness \citep{line16}, a quench pressure for nitrogen and carbon species \citep[e.g.][]{morley17}, and a power law haze opacity.  These models did not significantly improve the fit quality, and the main retrieval results were unchanged.  We also explored cases without the $30\times$ solar metallicity prior upper limit. Without this prior, metallicities up to $100\times$ were permitted, but other parameters were not significantly changed. 
%We also tested the effect of varying the planet's surface gravity within the published FIXME $\sigma$ range, and found that our results were again minimally changed.

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/retrieval.pdf}
\caption{Retrieved distributions for (a) metallicity, (b) carbon-to-oxygen ratio, (c) irradiation temperature, and (d) cloud-top pressure in bars. Black vertical lines show the median and $\pm1\,\sigma$ confidence interval (solid and dashed, respectively). Solar metallicity and C/O values are indicated with red vertical lines.}  \label{fig:retrieval}
\end{figure}

\subsubsection{Methane Abundance}
We ran two additional retrievals to explore whether methane depletion is the underlying cause of the inferred low C/O.  Even though water is the dominant absorber in the WFC3 bandpass, there is enough methane opacity to change the shape of the spectral features at a detectable level for high precision data.  In one test, we assumed chemical equilibrium but excluded methane opacity. This set-up resulted in a much broader distribution of C/O values ($0.02 - 1.6$ at $1\,\sigma$).

We also performed a ``free" retrieval that allowed the abundances of CH$_4$, H$_2$O and NH$_3$ to vary with no assumption of chemical equilibrium. The retrieved methane abundance is low ($3\times10^{-11} - 3\times10^{-6}$ at 68\% confidence). We rule out at $3\,\sigma$ confidence the predicted methane abundance for a composition with solar C/O and $3\,\times$ solar metallicity (the median metallicity from the chemical equilibrium retrieval).  Based on these tests, we conclude that the atmosphere of WASP-107b is depleted in methane relative to expectations for a solar abundance pattern. By contrast, the water abundance ($6\times10^{-6} - 2\times10^{-3}$) is consistent with predictions (FIXME).

\subsection{Condensate Properties}
In addition to the atmospheric retrieval, we also performed forward modeling of physically motivated, self-consistent clouds and hazes using the methods described in \cite{fortney08, morley15}.  We model clouds that form in cool atmospheres (Na$_2$S, KCl, ZnS; see \citealt{morley12}), for a range of metallicites from $1-50\times$ solar. We vary the cloud sedimentation efficiency from 3 to 0.1 (normal to highly lofted, small particulate clouds). None of the models produce sufficiently low amplitude spectral features to match the observed spectrum.  This result suggests that if clouds are muting WASP-107b's transmission spectrum, the mechanism for cloud particle formation and vertical lofting is very efficient and not captured by the modeling used here. Futher studies of cloud formation in very low gravity environments are required.

We also model an \emph{ad hoc} photochemical ``soot" layer near the top of the atmosphere. We predict the abundance of hydrocarbon haze precursors from previously published photochemical models for GJ 436b \citep{line11, morley17}. With the most efficient haze production (100\% of precursors form haze) and particle sizes around $0.03-0.1$ microns, the amplitude of the model water feature matches that of the observations.

These photochemical hazes are assumed to form from hydrocarbons generated by methane photolysis; however, we found in the previous section that methane is depleted.  Furthermore, even if the precursors were present, it is unclear whether the 100\% haze production efficiency is realistic.  However, organic hydrocarbons are not the only possible photolytic haze in cool H2-rich atmospheres. Sulfur chemistry may create additional haze material \citep{zahnle09, gao17}. Further modeling and laboratory work of the hazes that form in these conditions is needed for a satisfactory explanation of the data.

\section{Discussion and Conclusions} \label{sec:discuss}
We analyzed the atmospheric composition of WASP-107b based on retrieval of its near-infrared transmission spectrum and interior structure models of the planet's mass and radius.  Key results from this analysis include:
